A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2020-03-27 paper/paper - revision.Rnw
Edits to introduction, ready for Susan's comments

\mt{General paper to-do:}
\mt{1. Find a paper to cite the 95\% statistic}
\mt{Option: Shoe marks are most common at a crime scene (Rida et al., 2019)}
\mt{2. A zotero reference is missing details, remember to go find/fix it}
\mt{3. Cite papers with manual data collection}
\mt{4. Susan will send pictures of the prototype scanner from the grant proposal}
\mt{5. Argue in methods that we know real image quality will be lower but we needed to start somewhere}
\mt{6. Our goal is to narrow down what type of model works and assess the model(s) limitations}
\mt{7. Add to conclusion: Overall, this concept seems to work, but we need to modify the classificiation scheme. Thus, this work is publishable, because it needs work and community input. Need more research into human comphrehensible and yet useful classification schemes for this purpose. We need better class scheme, important and promising, getting this right matters}
\mt{8. Update abstract to reflect new framing of paper}
\mt{}

@@

From a statistical perspective, random match probability could theoretically be calculated by sampling a large number of shoes from a given population and counting the number of shoes in the sample whose tread patterns are similar to the given print. In practice, one potential method of sampling could be a low-profile scanner placed in a high traffic area. Such a scanner could take \mt{impressions or} pictures of shoes walking past, and those scans could then be used to create a database of shoes relevant to the local area.

However, studies by \fix{authors (yr)} and \fix{authors (yr)} demonstrate that even \fix{[a couple hundred]} shoe prints constitute an unweildy amount of data requiring \fix{[a large amount of data processing]}. Thus, in parallel with the development of such a scanner, it is necessary to develop a way to process large amounts of data automatically. \mt{The primary focus of this paper is to develop one such method for large-scale automatic data processing; in this case, "processing" refers to assessing} the similarity of the shoe tread patterns in the collected data. %\fix{Cut this?: This is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.}

Although footwear impressions are more commonly used in practice for footwear analysis than outsole images, early prototypes of a footwear scanner have suggested that cameras are more robust than pressure scanners for placement in public locations that may experience inclement weather or other non-controlled conditions. Thus, \fix{this is an image recognition problem}.

@@

There is precedent for using machine learning to identify tread patterns \mt{hunt through papers to find relevant references}

>>>>

\mt{General paper to-do:}
\mt{1. Find a paper to cite the 95\% statistic}
\mt{Option: Shoe marks are most common at a crime scene (Rida et al., 2019)}
\mt{2. A zotero reference is missing details, remember to go find/fix it}
\mt{3. Cite papers with manual data collection}
\mt{3.5. Incorporate recommended citations from reviewer 3}
\mt{4. Susan will send pictures of the prototype scanner from the grant proposal}
\mt{5. Argue in methods that we know real image quality will be lower but we needed to start somewhere}
\mt{6. Our goal is to narrow down what type of model works and assess the model(s) limitations}
\mt{7. Add to conclusion: Overall, this concept seems to work, but we need to modify the classificiation scheme. Thus, this work is publishable, because it needs work and community input. Need more research into human comphrehensible and yet useful classification schemes for this purpose. We need better class scheme, important and promising, getting this right matters}
\mt{8. Update abstract to reflect new framing of paper}
\mt{}

@@

From a statistical perspective, random match probability could theoretically be calculated by sampling a large number of shoes from a given population and calculating the proportion of shoes in the sample whose tread patterns are similar to the given print. In practice, one potential method of sampling could be a low-profile scanner placed in a high traffic area. Such a scanner could take \mt{impressions or} pictures of shoes walking past, and those scans could then be used to create a database of shoes relevant to the local area.

However, studies by \fix{authors (yr)} and \fix{authors (yr)} demonstrate that even \fix{[a couple hundred]} shoe prints constitute an unweildy amount of data requiring \fix{[a large amount of data processing]}. Thus, in parallel with the development of such a scanner, it is necessary to develop a way to process large amounts of data automatically. \mt{The primary focus of this paper is the development of one such method for large-scale automatic data processing, where "data processing" refers to assessing} the similarity of the shoe tread patterns in the collected data.

Although footwear impressions are more commonly used in practice for footwear analysis than outsole images, early prototypes of a footwear scanner have suggested that cameras are more robust than pressure scanners for placement in public locations that may experience inclement weather or other non-controlled conditions. Thus, \fix{we move forward by considering how to assess similarity of tread patterns in images, in particular.} More specifically, this assessment is done by identifying a set of relevant features and using image analysis methods to detect the presence or absence of the features in each image, and then defining similarity measures based on those features to determine how many shoes have a tread pattern similar to the pattern of interest.

@@

% From original intro:
In this paper, we discuss an alternate approach that uses an additional model layer to transfer the feature vector output into a vector of probabilities representing the detection of geometric elements in shoe tread images. Working within the feature space used by forensic examiners allows us to augment human-identified features with model output, assessing similarity of different shoe images on a feature set that is explicitly relevant to the domain. \fix{Ultimately, we can use the output probabilities to quantify similarity between tread patterns for the ultimate goal of calculating random match probability for a given tread pattern.}
